---
title: "p8105_hw6_kb2977"
author: "Francois Ban"
date: "11/18/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dbplyr)
```

###Problem 1:

```{r homicide_data}
homicide_data = read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")
```

####Part 1: Modifying and Cleaning Dataset
* Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake.

```{r cleaning}
homicide_data = homicide_data %>% 
  mutate(city_state = str_c(city, state, sep = ", ")) %>% 
  mutate(solved = disposition) %>%
  mutate(solved = recode(solved, 
                         "Closed without arrest" = "no", 
                         "Open/No arrest" = "no", 
                         "Closed by arrest" = "yes")) %>% 
  mutate(solved = as.factor(solved))
homicide_data = homicide_data[!(homicide_data$city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")), ]

```

* Modify victim_race to have categories white and non-white, with white as the reference category. Be sure that victim_age is numeric.

```{r modify victim_race}
homicide_data = 
  homicide_data %>%
  mutate(victim_race_cat = victim_race) %>%
  mutate(victim_race_cat = recode(victim_race_cat, 
                         "Hispanic" = "non-white", 
                         "Black" = "non-white", 
                         "Other" = "non-white",
                         "Asian" = "non-white")) %>% 
  mutate(victim_race_cat = fct_infreq(victim_race_cat)) %>% 
  mutate(victim_age = as.numeric(victim_age))
```

####Part 2: Baltimore, MD
* For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing black victims to white victims keeping all other variables fixed.

```{r fitted logistic regression for Baltimore}
logistic_balt = 
  homicide_data %>% 
  filter(city_state == "Baltimore, MD") %>% 
  mutate(victim_race = fct_relevel(victim_race, "White")) %>% 
  glm(solved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) %>% 
  broom::tidy() 

logistic_balt %>% 
  select(term, estimate, std.error) %>% 
  mutate(OR = exp(estimate)) %>% 
  mutate(lower_CI = exp(estimate - (1.96 * std.error)),
        upper_CI = exp(estimate + (1.96 * std.error))) %>% 
  filter(term == "victim_raceBlack") %>% 
  select(OR, lower_CI, upper_CI) %>% 
  knitr::kable(digits = 3)
```

####Part 3: GLM for Each City
* Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing black victims to white victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r ci for each city}
CI_for_each_city = function(each_city){

  CI = homicide_data %>% 
  filter(city_state == each_city) %>% 
  mutate(victim_race = fct_relevel(victim_race, "White")) %>% 
  glm(solved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  select(term, estimate, std.error) %>% 
  mutate(OR = exp(estimate)) %>% 
  mutate(lower_CI = exp(estimate - (1.96 * std.error)),
        upper_CI = exp(estimate + (1.96 * std.error))) %>% 
  filter(term == "victim_raceBlack") %>% 
  select(OR, lower_CI, upper_CI) 
  
CI
}
```

```{r nesting}
nest_glm_results =
  homicide_data %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(models = map(city_state, CI_for_each_city)) %>% 
  select(-data) %>% 
  unnest()

nest_glm_results %>% 
  knitr::kable(digits = 3)
```

####Part 4: Plot for Each City
*Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r OR plots}
OR_plot = nest_glm_results %>%
  ggplot(aes(x = city_state, y = OR)) + 
    geom_point(aes(reorder(city_state, OR)), size = 0.7, color = "BLUE") +
    geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), 
                  color = "RED", 
                  width = 0.1, 
                  alpha = 0.5) +
    labs(title = "Estimated Odds Ratio of Solved Homicides Comparing Black Victims to White Victimes in Each City",
        x = "City",
        y = "Estimated OR") + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 7))
OR_plot
```

*The odd ratios of solving homicides for black victims to white victims is highest for Tampa, FL and lowest for Boston, MA. Only about three cities (Birmingham, AL, Durham, NC, and Tampa, FL) have a odds ratio of greater than 1. The rest have an odds ratio of less than 1. Long Beach, CA seems to have an OR around 1 meaning the odds of solving homicides for black victims is similar, if not the same, to the odds of solving homicides for white victims in Long Beach, CA.

###Problem 2:

```{r birthweight_data}
birthweight_data = 
  read_csv(file = "./data/birthweight.csv") %>% 
  janitor::clean_names()
```

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.
